### 2024-07-08 rebuttals/rebuttals.tex

\end{mdframed}

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{orange} \textbf{Stefano}}, {\color{teal} \textbf{Stefan}} \& {\color{purple} \textbf{Sylvain}} - Many axes labels are missing, and the captions are often insufficient to read the figures. 
\end{mdframed}

\noindent I agree and have gone through all the figures of the manuscript to clarify them by adding axes labels or extending the caption where it was needed. Figures where y-axis (and its label) is missing are distribution (or density) plots (\ie Figures 3.6, 4.4, 4.5, 4.6, 4.8 and E.1).  Consequently, similarly to an histogram, the feature of interest and its span of values are represented on the x-axis. The y-axis has no real interest in terms of information. What matters is the shape of the curve to be able to point out for which values/ranges of values, the feature of interest occurred more or less often.  For this reason, I have not added y-axis for the aforementioned figures but rather improved their respective caption, emphasising, among other things, that these figures represent distribution of occurrences. For instance, the caption of {\color{blue}Figure 4.4} is now more explicit:

\begin{mdframed}[style=manuscript] % Comment from the reviewer
Exploration of the state space over the learning process: distribution of occurrence of cumulative emissions (left) and costs (right). Number of remaining attempts decrease with time since infeasible problem and solutions overshooting the \ce{CO2}-budget are discarded prematurely, \ie before reaching 2050. Besides infeasible problems, distributions labelled as ``Failure'' represent the attempts that overshot the \ce{CO2}-budget by 2050 at the latest. The majority of successful transitions have cumulative emissions much lower than the \ce{CO2}-budget and are cheaper than the REF case. 
\end{mdframed}

@@

\noindent Accounting for the depreciation of money with time (see previous comment), we considered that the assets are fully paid at the moment they are installed. For this reason, we consider the arithmetic average between the investment cost of the assets at the year starting the phase of installation and the year ending it. This choice was done for two main reasons: potential decommissioning and lack of data beyond 2050. In the Pathway version of EnergyScope, the asset will be used over its whole lifetime unless it is prematurely decommissioned. Therefore, annualising over the entire lifetime of the unit might induce a bias in case of this premature decommissioning. Then, for assets with longer lifetime and/or installed later in the transition, annualising over the lifetime would require to have data beyond 2050. Effort that I have not decided to make.

Regarding the impact on the results, my educated guess is that this impact would be limited as the technologies of a same sector are assumed to have similar trends in terms of the evolution with time of their respective investment costs. Consequently, the major impact would be to end up with a lower total transition costs but the design of the system would be marginally impacted.

Given these further explanations, I have not brought any modification to manuscript in this regard.

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{teal} \textbf{Stefan}} - Interest rate must be revisited if a paper is published.
\end{mdframed}

\noindent See previous comments.

@@

\subsubsection{Power grid and transport infrastructure}

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{purple} \textbf{Sylvain}} - How do you model the power grid interconnections with the neighbouring countries as well as the transport network of the different energy carriers within the country? p.63 - The model decides to use ammonia instead of SNG $\rightarrow$ is the transport infrastructure accounted for? p.73 - How is the transport infrastructure accounted for?
\end{mdframed}

\noindent Regarding the power grid interconnections with the neighbouring countries, similarly to the import of other energy carriers, there is no representation of those physical routes. The whole-energy system of the country is represented as a single-node and the imported energy commodities are available ``at the front door'' of the system. It goes the same way for the infrastructures to transport energy carriers within the system. As it is represented assuming a copper-plate approach, all the demands/resources/conversion technologies are concentrated in one single node. This is now explicitly written {\color{blue}in a new paragraph of Section 2.2}:

\begin{mdframed}[style=manuscript] % Modification brought to the manuscript 
Regarding the power grid interconnections with the neighbouring countries, similarly to the import of other energy carriers, there is no representation of those physical routes. The imported energy commodities are available ``at the front door'' of the system.  In terms of spatial resolution of the country itself, it is modelled as a single-node. This is similar to the copper plate model of a power grid. In other words, the infrastructures to transport the energy carriers within the country are not considered.  It is assumed that the demands have to be supplied by the production, regardless of the flows between the producers and the consumers. Yet, adapting the networks is accounted for in terms of the required investments. For instance, a high share of VRES requires an investment to reinforce the power grid (\ie 368\,M€/GW of additional installed capacity of VRES). 
\end{mdframed}

@@

\noindent As discussed during the private defense, similarly to EnergyVille \cite{PATHS2050}, we have considered the same CAPEX for SMR as the one considered for conventional nuclear power plants. Where EnergyVille considers a nominal CAPEX of 7500€/kW (like the current large Gen III design of Hinkley Point in the UK), we considered 4850€/kW like the CAPEX of conventional nuclear units in EnergyScope.

On top of the uncertainty applied to the CAPEX of SMR [-40\%, 44\%] considered in the thesis, I carried out a test on these assumptions by plotting the LCOE of electricity-production units considering different assumptions for SMR (see Figure \ref{fig:LCOE_comparison}). Even considering 50\% more expensive SMR subject to a higher discount rate (5\% versus 1.5\% for the other technologies), we notice that SMR keeps on being more cost-competitive than other flexible production units, like e-ammonia CCGTs, that are substituted by SMR when the latter can be installed (see Chapter 3). In other words, it means that the optimistic values assumed for SMR in the thesis lead to a cheaper whole-energy system but does not affect its design, \ie the installed technologies.  Consequently, no further modification has been brought to the thesis regarding this comment.

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.49\textwidth]{LCOE_line_2.pdf}
\includegraphics[width=0.49\textwidth]{LCOE_line_3.pdf}
\label{fig:LCOE_comparison}
\caption{LCOE per technology producing exclusively electricity considered in the thesis (left) and same LCOE but, this time, with a higher CAPEX (7270€/kW versus 4850€/kW in the thesis and a discount rate of 5\% on this technology. Even considering these less favourable assumptions for SMR, it still outcompetes other flexible production units like CCGT running on e-ammonia.}
\end{figure}

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{teal} \textbf{Stefan}} - SMR must be revisited if a paper is published.
\end{mdframed}

\noindent See previous comments.

@@

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{orange} \textbf{Stefano}} - p. 70: ``in other words, the variation of the parameter $f_{\mathrm{max,SMR}}$…''. I am not sure that this sentence is equivalent to the previous one. Please clarify.
\end{mdframed}

\noindent The Sobol' index is the ratio between the variance of the total transition cost due to the variation of one parameter and the total variance of the total transition cost. If $f_{\mathrm{max,SMR}}$ takes a value between 0 and 0.59, the impact on the system, and, consequently, on its total transition cost, is the same: no SMR installed. A wide range of variation of the parameter leading to zero variation of the output of interest is one of the reasons to explain why the Sobol' index of this parameter is low. To clarify this point, I have rephrased the sentence {\color{blue}in Section 3.2.1}:

\begin{mdframed}[style=manuscript] % Modification brought to the manuscript
Given the uncertainty characterisation presented in Section 2.4, there are 60\% chance that no SMR could be installed. \textbf{When $f_{\mathrm{max,SMR}}$ takes a value between 0 and 0.59, the variation of the total transition cost is only due to the variation of the other parameters. }
\end{mdframed}

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{orange} \textbf{Stefano}} - p. 71, Table 3.1: isn’t the low impact simply due to the low assumed potential? Is this finding meaningful?

\noindent {\color{purple} \textbf{Sylvain}} - p. 70 - But that's because the CAPEX is way too low. With more realistic values, the Sobol would be much higher.
\end{mdframed}

\noindent These assumed low potential and CAPEX are indeed some reasons why the SMR-related uncertain parameters have a limited impact on the variation of the total transition cost as already stated in the first paragraph of Section 3.2.1: ``However, given its limited deployment, up to 6\,GW, SMR-related parameters have a lower impact on the variation of the total transition cost.''. However, this is not the only reasons as detailed earlier in the same paragraph: the late potential availability (2040 at the soonest) as well as the ``cheap and low-emitting fuel, the long lifetime leading to lower annualised CAPEX and higher salvage value.''.

@@

\noindent Indeed, the efficiency only has an impact on the amount of primary uranium needed to produce the electricity. Consequently, I have removed this part of the sentence {\color{blue}in the first paragraph of Section 3.1.1}.

\begin{mdframed}[style=manuscript] % Modification brought to the manuscript 
Overall, given the restriction on yearly availability and the slightly higher electrification [...]
\end{mdframed}

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{purple} \textbf{Sylvain}} - p. 64 - 15 GW of PV in 2025. Realistic?
\end{mdframed}

\noindent In the continuity of Gauthier Limpens' thesis  and in line with the rest of the research field, the objective of this thesis is to ``illustrate the priorities to be taken in each of the sectors and the efforts to make in one sector to facilitate the integration of certain technologies into other sectors'' \cite{limpens2021generating}. This work highlights the ``must-do'' or ``no-regret investments'' to respect the (very) ambitious \ce{CO2} budget imposed to the model. In this sense, besides the maximum potential of 59\,GW solar-PV, we have not made assumption on the rate of deployment. 15\,GW of PV in 2025 can be considered as very optimistic and Elia forecasts to have these 15\,GW by 2030 \cite{Elia_2024_2034}. I agree that not limiting the rate of deployment of PV (and wind turbines) is a strong assumption. For this reason, in the perspectives of this work (see Conclusion), I have already pointed this out as ``a future work would be to constrain the deployment of local renewables.

Given these explanations, I have not brought further modification to the manuscript in this regard.

@@

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{purple} \textbf{Sylvain}} - p. 67 - It really depends on what is considered primary energy in the case of nuclear
\end{mdframed}

\noindent As specified in the table about SMR features (see previous comment), the considered fuel for SMR is uranium, as it is the case for conventional nuclear power plants. Given this, I have not brought modification to the manuscript in this regard.

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{purple} \textbf{Sylvain}} - p. 69 - No curtailment at all? It would be nice to have the exact number.
\end{mdframed}

\noindent There is indeed some curtailment, which remains very limited compared to the total production of solar-PV: 0.3\,TWh versus 61\,TWh. This information has been added {\color{blue}in the paragraph dedicated to ``High-temperature heat'' of Section 3.1.3}:

\begin{mdframed}[style=manuscript] % Modification brought to the manuscript 
In the REF case, industrial electric heaters are mainly used to absorb the ``over-production'' of the 59\,GW solar-PV when fully deployed, in the sunny days, and limit the yearly curtailment to a maximum of 0.3\,TWh, 0.5\% of the yearly 61\,TWh produced by solar-PV.
\end{mdframed}

@@

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{purple} \textbf{Sylvain}} - p. 69 - Very strange. Should be the same as for industrial boilers. No smart charging?
\end{mdframed}

\noindent The difference between BEV and industrial electric heaters is that BEV is already a preferred option for electrification, even the REF case. BEV completely substitute diesel and gasoline-cars as soon as possible in the transition and there cannot be more cars as private mobility is limited to supply up to 50\% of the demand of passengers mobility. Unlike high-temperature heat, there is no modal swap due to the installation of SMR.

Given these explanations, I have not brought further modification to the manuscript in this regard.

@@

\noindent As written in Chapter 3: ``In practice, the discount rate would vary depending on the technology investment risk''. Depending on the technology readiness level, private investors would be more or less risk-averse. However, EnergyScope only considers a vision of a central-planner where a single agent makes the investment decisions without differentiating the sources of the capitals, \ie private and public. Assuming a differentiated discount rate would require to make assumptions about the private-public distribution of the capitals. These further explanations for keeping a single value of discount rate have been added {\color{blue} after Eq. (1.3) when discussing the annualised phase factor}.

\begin{mdframed}[style=manuscript] % Modification brought to the manuscript 
The discount rate accounted for in the annualisation factor is considered as identical for all the technologies of the system. In practice, the discount rate would vary depending on the technology investment risk. Depending on the technology readiness level, private investors would be more or less risk-averse. However, EnergyScope only considers the vision of a central-planner where a single agent makes the investment decisions without differentiating the sources of the capitals, \ie private and public. Assuming a differentiated discount rate would require to make assumptions about the private-public distribution of the capitals. For this reason, we have decided to keep an identical value of discount rate for all the technologies.
\end{mdframed}

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{purple} \textbf{Sylvain}} - p. 72 - Battery trucks are gaining momentum. Is the hydrogen route really the way to go?
\end{mdframed}

\noindent In general, the model gives fuel-cell trucks as the favoured option for the future of road freight transport given more cost-competitive options for electrification like private mobility and low-temperature heat. However, as illustrated in the ROB roadmap (see Section 5.2.1), battery trucks seem to be the cost-optimal solution to transition between diesel and either hydrogen trucks, when fossil and electrofuels are assumed to be expensive.

In the Global Sensitivity Analysis, electric trucks are the favoured option for a limited set of samples, about 10\% (see Appendix D.3).

Given these explanations and what is already present in the text, I have not brought further modification to the manuscript in this regard.

@@

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{purple} \textbf{Sylvain}} - p. 78 - Why this threshold? Shouldn't it follow the potential capacity?
\end{mdframed}

\noindent The threshold locates at 60\% of the uncertain parameter $f_{\mathrm{max,SMR}}$ which represents the potential capacity of SMR. As presented in Section 2.4, this threshold corresponds to the possibility to install up to 6\,GW of SMR in 2050. As already written in Section 3.2.2: ``For the installed capacity of SMR, it is the availability of the technology that drives its installation (Figure 3.12). Not shown here but all the samples of the GSA highlight that SMR is installed to its maximum capacity, \ie 6GW, as soon as possible.'' To smooth the trend curves of the different uncertain parameters, I have applied a rolling average.  For this reason, the 60\%-threshold does not appear on the Figure as sharp as it actually is.

Given these explanations and what is already present in the text, I have not brought further modification to the manuscript in this regard.

@@

\begin{mdframed}[style=manuscript] % Modification brought to the manuscript

\end{mdframed}

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{orange} \textbf{Stefano}} - p. 94: Figure 4.6 seems to be less revealing than other figures. Also the figure caption seems to contradict the first statement following the figure. It would be important to clarify: which are the actions that emerge to actually matter?
\end{mdframed}

\noindent Indeed, Figure 4.6 is less revealing as it shows no big trend in terms of neither impacting actions nor their timing in the transition.  It is detailed in the first paragraph of Section 4.2.3 after which the figure should appear in the manuscript. I have worked to make Figures as close as possible to their related paragraph. However, in some cases, several other paragraphs come between them. To identify the actions that emerge to actually matter, we need to check if they are binding or not, as detailed in the subsequent paragraphs and in Figure 4.7.

Given these explanations and what is already present in the text, I have not brought further modification to the manuscript in this regard, except minor adjustment to get Figure 4.6 as close as possible to its related paragraph.

@@

\begin{mdframed}[style=manuscript] % Modification brought to the manuscript

\end{mdframed}

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{purple} \textbf{Sylvain}} - p. 87 - What is a step?
\end{mdframed}

\noindent In the context of Reinforcement Learning, a step consists in taking an action, collecting a reward based on the previous state and the taken action and ending in a new state. Even though it was already mentioned in Chapter 1, the fact that the results come further from the methodology part makes it harder for the reader to remember the meaning of a step. It is now more explicitly defined {\color{blue}in the first paragraph of Section 1.3.2}:

\begin{mdframed}[style=manuscript] % Modification brought to the manuscript 
From this state, the agent takes a step: action, reward, new state.
\end{mdframed}

\noindent and reminded {\color{blue}in the first sentence of Section 4.2.1}:

\begin{mdframed}[style=manuscript] % Modification brought to the manuscript 
The learning phase has been split into batches of 500 steps,\ie 500 sequences of state-action-reward-new state. 
\end{mdframed}

@@

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{purple} \textbf{Sylvain}} - p. 96 - The early constraints on generation avoid lock-in situations correct?
\end{mdframed}

\noindent Correct. This confirms what \citet{vogt2018starting} detailed in their works. To emphasise on this point, I have added one sentence {\color{blue}in Section 4.2.3}:

\begin{mdframed}[style=manuscript] % Modification brought to the manuscript 
In line with \citet{vogt2018starting}, the early constraints on generation and the sharper decrease of the emissions avoid lock-in situations.
\end{mdframed}

@@

\begin{mdframed}[style=manuscript] % Modification brought to the manuscript

\end{mdframed}

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{purple} \textbf{Sylvain}} - p. 100 - What/where are the no-go zones? Is there a narrative to describe them ?
\end{mdframed}

\noindent No-go zones are introduced in the second paragraph of Chapter 4: ``[...] as well as no-go zones where the chances of succeeding the transition, i.e., respecting the \ce{CO2}-budget, are very limited.'' It is then detailed at the end of Section 4.2.2: ``Below the near-term threshold of $\sim$60\%, this is the ``no-go zone'' where succeeding the transition becomes unlikely, except
if betting on the future installation of SMR. '' and in the caption of the related figure. Given the information already present in the manuscript, no modification has been brought in this regard.

@@

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{violet} \textbf{Christophe}} - The concept of binding constraint was not clear.
\end{mdframed}

\noindent
Besides the information given in Section 4.2.3 of the manuscript, I do not see further explanations that could clarify more the concept of binding constraint in a Linear Programming (LP) problem. Consequently, regarding this comment, there has not been further modification brought to the manuscript.

@@

{\color{blue} End of second paragraph of section 1.4.1}:

### 2024-07-08 rebuttals/rebuttals.tex

\begin{mdframed}[style=manuscript] % Modification brought to the manuscript
Moreover, this means that \textbf{the coefficient} $\alpha_{ki}$, \ie the component of $\bm{\alpha}_{\mathbf{k}}$ related to the $i^{\text{th}}$ original variable, $x_i$,  gives its weight in the $k^{\text{th}}$ PC, \ie $z_k$. 
\end{mdframed}

\section{Conclusion}
\label{Conclusion}

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{orange} \textbf{Stefano}} \& {\color{purple} \textbf{Sylvain}} - p. 119: the statement mentioning the ``major added value of the perfect foresight'' approach seems surprising and needs to be clarified.  The decreasing \ce{CO2} emissions trajectories shouldn't be a S-shaped (logistic) curve?
\end{mdframed}

\noindent I agree that this sentence and what came next focused on the perfect foresight approach whereas the rest of the paragraph rather targets the RL-based myopic results. Therefore, I have reformulated {\color{blue}the end of this paragraph} and focused on the shape of the decreasing \ce{CO2} trajectories. These are indeed S-shaped curves but that are sharper for the myopic approach than in the PF results.

\begin{mdframed}[style=manuscript] % Modification brought to the manuscript
This work also highlighted a major difference between the perfect foresight (PF) approach and the agent-based myopic transitions. Given the assumed full knowledge on the future, the trajectories to decrease the \ce{CO2} emissions resulting from perfect foresight are more progressive, and foreseen to be more applicable, compared to a sharper drop at the early stages of the transition, like in the myopic approach.
\end{mdframed}

\begin{mdframed}[style=comment] % Comment from the reviewer
{\color{teal} \textbf{Stefan}} - The conclusion is a summary and should rather be a discussion.
\end{mdframed}

\noindent I decided to split the conclusion into four blocks of paragraphs. The first two (from page 117 to the top of page 120, in the version of the manuscript submitted before the private defense) are indeed a summary of (1) what we have done in this thesis (case study and methodology) and (2) the results we have got. The two last blocks, however, take a step back from the results and aim at drawing main conclusions, discussing them and providing decision-makers with further insightful information and methods to plan the energy transition. Consequently, I have not brought modification to the manuscript in this regard.

### 2024-07-08 appendices/app_EnergyScope.tex
Go, go, go...

In Equation (eq:salvage_app), $\tau_{phase}$ is the annualisation factor corresponding to the phase where the technology was built. With this expression, if a technology is still operational for half its lifetime after 2050, then half its initial investment is subtracted to the total transition cost. However, one could wonder if a technology that is still operational for half of its lifetime is not worth less than half of its initial investment cost, given the depreciation of the asset. In that case, the annualisation factor should account for the last phase of the optimisation (\ie 2045\_2050) when the residual value of the technologies are calculated and subtracted from the total transition cost.

On top of this, we also investigated the impact of accounting for the decommissioned technologies or not in the expression of the salvage value.  In Equation (eq:salvage_app), if a technology is decommissioned before 2050, then the salvage value of this technology will be 0. This is relevant since a technology decommissioned before 2050 would not be available after 2050. The issue is that the model could keep unnecessary technologies to subtract their salvage value from the total transition cost in order to decrease it. This would be done if the fixed OPEX of an unused technology is lower than its salvage value. Therefore, one could wonder if it would not be better to take into account the decommissioned technologies in the salvage value. In that case, a technology that would have been operational after 2050 but that had been prematurely decommissioned before 2050, would have a salvage value.

>>>>

In Equation (\ref{eq:salvage_app}), $\tau_{phase}$ is the annualisation factor corresponding to the phase where the technology was built. With this expression, if a technology is still operational for half its lifetime after 2050, then half its initial investment is subtracted to the total transition cost. However, one could wonder if a technology that is still operational for half of its lifetime is not worth less than half of its initial investment cost, given the depreciation of the asset. In that case, the annualisation factor should account for the last phase of the optimisation (\ie 2045\_2050) when the residual value of the technologies are calculated and subtracted from the total transition cost.

On top of this, we also investigated the impact of accounting for the decommissioned technologies or not in the expression of the salvage value.  In Equation (\ref{eq:salvage_app}), if a technology is decommissioned before 2050, then the salvage value of this technology will be 0. This is relevant since a technology decommissioned before 2050 would not be available after 2050. The issue is that the model could keep unnecessary technologies to subtract their salvage value from the total transition cost in order to decrease it. This would be done if the fixed OPEX of an unused technology is lower than its salvage value. Therefore, one could wonder if it would not be better to take into account the decommissioned technologies in the salvage value. In that case, a technology that would have been operational after 2050 but that had been prematurely decommissioned before 2050, would have a salvage value.

### 2024-07-08 chap_RL/chap_RL.tex
Go, go, go...

The learning phase has been split into batches of 500 steps. %For the first 100 steps of each batch, the \gls{RL}-model collects transitions before learning starts. This makes sure replay buffer is full enough for useful updates. 
At the end of each batch, the up-to-date policy, \ie the \gls{NN}, is saved. This way, we can assess the progress in the learning process and its convergence (see Figure \ref{fig:reward_success}).  The mean reward increases rapidly at the beginning of the learning process before reaching a plateau where the optimisation of the policy becomes more marginal.  As this successes drive indirectly the agent's optimisation, it shows that the reward function (see Figure \ref{fig:Reward}) leads towards more and more successes. However, given the wide range of uncertainty of some parameters and the agent's levers of action, this success rate stays limited at the end of the learning process. In other words, there are conditions where it is impossible for the agent to succeed the transition.

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{Cum_gwp_cost_3.pdf}
\caption{Exploration of the state space over the learning process: distribution of occurrence of cumulative emissions (left) and costs (right). Number of remaining attempts decrease with time since infeasible problem and solutions overshooting the \ce{CO2}-budget are discarded prematurely, \ie before reaching 2050. Distributions labelled as ``Failure'' represent the attempts that overshot the \ce{CO2}-budget by 2050 at the latest. The majority of successful transitions have cumulative emissions much lower than the \ce{CO2}-budget and are cheaper than the REF case. }
\label{fig:Cum_gwp_cost}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{RE_in_mix_Efficiency.pdf}
\caption{Exploration of the state space over the learning process: distribution of occurrence of share of renewable energy carriers in the primary energy mix (left) and efficiency (right).  Number of remaining attempts decrease with time since infeasible problem and solutions overshooting the \ce{CO2}-budget are discarded prematurely, \ie before reaching 2050. Distributions labelled as ``Failure'' represent the attempts that overshot the \ce{CO2}-budget by 2050 at the latest. Integration of local \gls{VRES} at early stages then massive import of electrofuels later are needed to secure successful transitions. Below a near-term threshold ($\sim$60\%), the chances of success are limited, \ie no-go zones. Efficiency is a less valuable information for the agent to succeed transitions as failures and successes indistinguishably spread over the whole range.}
\label{fig:RE_in_mix_Efficiency}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{Actions_learning.pdf}
\caption{Exploration of the action space over the learning process: distribution of occurrence of the different actions at the time they are taken by the agent. Number of remaining attempts decrease with time since infeasible problem and solutions overshooting the \ce{CO2}-budget are discarded prematurely, \ie before reaching 2050. Distributions labelled as ``Failure'' represent the attempts that overshot the \ce{CO2}-budget by 2050 at the latest. Besides this wide exploration, successes and failures indistinguishably spread over the whole ranges of actions. In other words, it does not seem to be any clear set of actions to support successful transitions.}
\label{fig:Actions_learning}
\end{figure}

@@

After filtering out failures of the learning episodes and keeping only the successful transitions, only a limited set of the actions are binding and have an actual impact on the result of the optimisation in EnergyScope Pathway (see Figure \ref{fig:Binding_learning}). This allows identifying key actions to support the myopic transition. Limiting the \gls{GWP} in the near-term is a key-factor for success. However, this action has a binding effect on the environment only at the end of the transition. The range over which limiting the use of fossil gas binds the optimisation is wider. Compared to other non-renewable fuels, this is due to the longer use of this energy carrier favoured by its low \gls{GWP} (the second after uranium) and its versatility (applications in the electricity, heat and mobility sectors).

>>>>

The learning phase has been split into batches of 500 steps,\ie 500 sequences of state-action-reward-new state. %For the first 100 steps of each batch, the \gls{RL}-model collects transitions before learning starts. This makes sure replay buffer is full enough for useful updates. 
At the end of each batch, the up-to-date policy, \ie the \gls{NN}, is saved. This way, we can assess the progress in the learning process and its convergence (see Figure \ref{fig:reward_success}).  The mean reward increases rapidly at the beginning of the learning process before reaching a plateau where the optimisation of the policy becomes more marginal.  As this successes drive indirectly the agent's optimisation, it shows that the reward function (see Figure \ref{fig:Reward}) leads towards more and more successes. However, given the wide range of uncertainty of some parameters and the agent's levers of action, this success rate stays limited at the end of the learning process. In other words, there are conditions where it is impossible for the agent to succeed the transition.

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{Cum_gwp_cost_3.pdf}
\caption{Exploration of the state space over the learning process: distribution of occurrence of cumulative emissions (left) and costs (right). Number of remaining attempts decrease with time since infeasible problem and solutions overshooting the \ce{CO2}-budget are discarded prematurely, \ie before reaching 2050. Besides infeasible problems, distributions labelled as ``Failure'' represent the attempts that overshot the \ce{CO2}-budget by 2050 at the latest. The majority of successful transitions have cumulative emissions much lower than the \ce{CO2}-budget and are cheaper than the REF case. }
\label{fig:Cum_gwp_cost}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{RE_in_mix_Efficiency.pdf}
\caption{Exploration of the state space over the learning process: distribution of occurrence of share of renewable energy carriers in the primary energy mix (left) and efficiency (right).  Number of remaining attempts decrease with time since infeasible problem and solutions overshooting the \ce{CO2}-budget are discarded prematurely, \ie before reaching 2050. Besides infeasible problems, distributions labelled as ``Failure'' represent the attempts that overshot the \ce{CO2}-budget by 2050 at the latest. Integration of local \gls{VRES} at early stages then massive import of electrofuels later are needed to secure successful transitions. Below a near-term threshold ($\sim$60\%), the chances of success are limited, \ie no-go zones. Efficiency is a less valuable information for the agent to succeed transitions as failures and successes indistinguishably spread over the whole range.}
\label{fig:RE_in_mix_Efficiency}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\textwidth]{Actions_learning.pdf}
\caption{Exploration of the action space over the learning process: distribution of occurrence of the different actions at the time they are taken by the agent. Number of remaining attempts decrease with time since infeasible problem and solutions overshooting the \ce{CO2}-budget are discarded prematurely, \ie before reaching 2050. Besides infeasible problems, distributions labelled as ``Failure'' represent the attempts that overshot the \ce{CO2}-budget by 2050 at the latest. Besides this wide exploration, successes and failures indistinguishably spread over the whole ranges of actions. In other words, it does not seem to be any clear set of actions to support successful transitions.}
\label{fig:Actions_learning}
\end{figure}

@@

After filtering out failures of the learning episodes and keeping only the successful transitions, only a limited set of the actions are binding and have an actual impact on the result of the optimisation in EnergyScope Pathway (see Figure \ref{fig:Binding_learning}). This allows identifying key actions to support the myopic transition. Limiting the \gls{GWP} in the near-term is a key-factor for success. However, this action has a binding effect on the environment only at the end of the transition. The range over which limiting the use of fossil gas binds the optimisation is wider. Compared to other non-renewable fuels, this is due to the longer use of this energy carrier favoured by its low \gls{GWP} (the second after uranium) and its versatility (applications in the electricity, heat and mobility sectors). In line with \citet{vogt2018starting}, the early constraints on generation and the sharper decrease of the emissions avoid lock-in situations.

### 2024-07-08 chap_atom_mol/chap_atom_mol.tex
Go, go, go...

The main impact of including \gls{SMR} from 2040 onward on the high-temperature heat sector is (i) its higher direct electrification and (ii) the reduction of more efficient heat-and-power co-generation in the benefit of single-output industrial boilers. In the REF case, industrial electric heaters are mainly used to absorb, instead of curtailing, the ``over-production'' of the 59\,GW solar-PV when fully deployed, in the sunny days. With \gls{SMR}, by 2050, an additional 1.7\,GW (+13\%) of these heaters can rely on a more constant supply of defossilised electricity. This increases the load factor (31\%) and yearly production (48\%) of these industrial electric heaters. Then, given the 44.6\,TWh of electricity produced by \gls{SMR}, cogeneration units are less relevant and, by 2050, 2.6\,GW industrial gas boilers completely substitute \gls{CHP} to produce 16.6\,TWh (\ie 23\%) of the total production of high-temperature heat.\\

>>>>

The main impact of including \gls{SMR} from 2040 onward on the high-temperature heat sector is (i) its higher direct electrification and (ii) the reduction of more efficient heat-and-power co-generation in the benefit of single-output industrial boilers. In the REF case, industrial electric heaters are mainly used to absorb the ``over-production'' of the 59\,GW solar-PV when fully deployed, in the sunny days, and limit the yearly curtailment to a maximum of 0.3\,TWh, 0.5\% of the yearly 61\,TWh produced by solar-PV. With \gls{SMR}, by 2050, an additional 1.7\,GW (+13\%) of these heaters can rely on a more constant supply of defossilised electricity. This increases the load factor (31\%) and yearly production (48\%) of these industrial electric heaters. Then, given the 44.6\,TWh of electricity produced by \gls{SMR}, cogeneration units are less relevant and, by 2050, 2.6\,GW industrial gas boilers completely substitute \gls{CHP} to produce 16.6\,TWh (\ie 23\%) of the total production of high-temperature heat.\\

### 2024-07-08 chap_case_study/chap_case_study.tex
Go, go, go...

@@

\section{Conversion technologies}
\label{sec:cs:technologies}
As the end-use demands are defined as energy (and non-energy with the \gls{NED}) services rather than a certain quantity of oil or solar irradiance, for instance, technologies are implemented to convert these resources into the end-use demands. Besides their CAPEX, OPEX and lifetime defined in Section \ref{sec:meth:ES}, production and conversion technologies (\ie \gls{CCGT}, car or boiler) have a conversion efficiency whereas storage technologies (\ie thermal storage, battery or molecule storage) have charge/discharge losses. There are also infrastructure technologies. They encompass, for instance, the power grid, the \gls{DHN} or technologies to produce intermediate energy carriers (\eg wood pyrolysis, biomethanolation or steam methane reforming to produce hydrogen). Not digging into too much details about the exhaustive list of these technologies presented in a previous work \cite{limpens2021generating}, this section rather focuses on the implementation of \glsxtrfull{SMR} and the technologies to supply the \glsxtrfull{NED}.

>>>>

Regarding the power grid interconnections with the neighbouring countries, similarly to the import of other energy carriers, there is no representation of those physical routes. The imported energy commodities are available ``at the front door'' of the system.  In terms of spatial resolution of the country itself, it is modelled as a single-node. This is similar to the copper plate model of a power grid. In other words, the infrastructures to transport the energy carriers within the country are not considered.  It is assumed that the demands have to be supplied by the production, regardless of the flows between the producers and the consumers. Yet, adapting the networks is accounted for in terms of the required investments. For instance, a high share of VRES requires an investment to reinforce the power grid (\ie 368\,M€/GW of additional installed capacity of \gls{VRES}).

@@

\section{Conversion technologies}
\label{sec:cs:technologies}
As the end-use demands are defined as energy (and non-energy with the \gls{NED}) services rather than a certain quantity of oil or solar irradiance, for instance, technologies are implemented to convert these resources into the end-use demands. Besides their CAPEX, OPEX and lifetime defined in Section \ref{sec:meth:ES}, production and conversion technologies (\ie \gls{CCGT}, car or boiler) have a conversion efficiency whereas storage technologies (\ie thermal storage, battery or molecule storage) have charge/discharge losses. There are also infrastructure technologies. They encompass, for instance, the power grid, the \gls{DHN} or technologies to produce intermediate energy carriers (\eg wood pyrolysis, biomethanolation or steam methane reforming to produce hydrogen).

Not digging into too much more details about the exhaustive list of these technologies presented in a previous work \cite{limpens2021generating}, this section rather focuses on the implementation of \glsxtrfull{SMR} and the technologies to supply the \glsxtrfull{NED}.
